The Lack of Accecibility to Media Outlets in Tanzinia in 1996

### 2022-04-11 outputs/paper/paper.Rmd
Final Commit

---
title: "The Lack of Accecibility to Media Outlets in Tanzinia in 1996 May Explain Absences of Benefits Seen Through Mass Media"
subtitle: "Data largely uses data on females "
author: 
  - Oluwabusayomi Adekuajo
thanks: "Code and data are available at: https://github.com/rubyzero10/Analysis_of_Mass_Media.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The significance of communication on the development of a country cannot be overestimated as the ability to spread information to a large audience supports community and its progression. Lack of progression in developing countries raises the concern of whether beneficial effects of mass media can be observed.  In this paper the prevalence of mass media outlets in Tanzania in 1996 was determined, in which most individuals in rural did not have access to any media sources included. Our findings have implications to develop affordable access to media through public spaces."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

@@

Communication, more than ever, has great significance to civilisation, so much so that is necessary for survival. The constant development of technology has led to the progression of media to what is known as mass media. Its key function is to communicate various messages to a large audience; common outlets include radio, newspapers, television, internet, social media, and more. The ability to bring information to the masses concerns the successiveness of a society, as some communities have the resources, such as easy access to internet, to take advantage of these communication outlets more than others that would face struggles to obtain the same source. As the strength of cooperation, obtained through utilization of mass media, is necessary for the development of a community, after the late 90s, New print, radio, and TV enterprises were established, from commercial media outlets to church and community-based stations, in east African developing countries including Tanzania(@spurk2017need). The benefits of this new media was predicted to promote democratic support and enhance, accountability and participation of the public sector (@spurk2017need).

Proper implementation of mass media was expected to have led to numerous positive outcomes, however, fulfilment of such expectation had yet to be met despite its investment in east Africa. In this paper, to determine that people are taking advantage of available mass media, we will analyse the proportion of mass media in the different regions of Tanzania. We compare the prevalence of all media outlets, along with discover possible preferences of media outlets. Media use was further compared by gender to visualize possible influence of gender roles on access to media. With this information we tried to determine circumstances in which benefits of media could be seen in developing countries, specifically Tanzania. Findings show that there was a large percentage of both men and women, in multiples regions of Tanzania, that did not use any forms of mass media. Out of those that that did use any or all forms of mass media, the most common use of was radio when compared to reading the newspaper or watching television. Findings are in line with past research that determined that the development of radio in Tanzania has been far more rapid than that of the press, with the potential of reaching the whole country (@probs). Radio is the most common source of media in developing countries due to its affordability and use of less electricity. Thus our finings imply that in order to achieve the right environment to support the development of media, economic viability must be considered and optimized.

The remainder of the paper is split into four sections. Section \@ref(data) explains the data source and collection methodology, potential bias and ethics issues, as well as our selected data’s characteristics. Section \@ref(results) presents the findings from our methodology. Section \@ref(discussion) provides discussion of results, the weaknesses as well as next steps to expand on the findings of the report. Section \@ref(datasheet) includes data sheet for the dataset put together.

@@

The 1996 Tanzania Demographic and Health Survey is a nationally representative survey with the main purpose to provide information on fertility, family planning, infant and child mortality, maternal and child health and nutrition, knowledge and attitudes of AIDS, and female circumcision(@tanzania). The TDHS is the third national samples survey of it its kind, practicing surveys since 1991. The general objected of the 1996 TDHS, which concerned the data used, is national-level data that will allow the calculation of demographic rates, particularly fertility and childhood mortality rates(@tanzania). The purpose of the TDHS was to provide estimated to the whole country including rural and urban areas.

@@

After the sampling process describes above, 8900 households were selected in which 8141 were occupied as the target sample size. 7969 individuals were interviewed showing a response rate of 98 percent. Out of the eligible 8900 households, 8501 was the target population for females age 15- 49, in which 96 percent were actually interviewed. In contracts, only 2658 was the target population for men age 15-59, from that  85 percent were interviewed. The TDHS thus collected data from 8,120 women aged 15-49 and 2,256 men aged 15-59 in Tanzania in 1996.

## Strength

The field staff conducting the interviews had to undergo intensive training consisting of interviewing techniques, detailed review of questionnaire items, mock interviews and much more, before being selected for the fieldwork. A series of assessments were given and graded from selected interviews and those who shows extra understood and had the ability to detect error in completed questionnaires were later chose to be editors. This training period consisted over the course of three weeks.

@@

In urban areas one member households were more common, as they were more likely to lock up their houses for safety. This made interviewing more difficult and response rate lower in such areas. Response rate was also lower among men despite the larger range in age and lower target sample size. This was observed due to lower and more frequent absence of men seen in household at time of interview, due to work status. This fact must be considered when comparing results between men and women.

@@

Interested in understanding the prevalence of mass media outlets in Tanzania households, data showing percentage of women and men who usually read the newspaper once a week, watch television once a week or listen to radio weekly in Tanzania 1996 was selected. The full table can be seen in Figure \@ref(fig:image), in which specifically region data was selected for the purposes of this paper. Data was gathered as PDF and read and manipulated through use of pdftools @pdf and tidyverse @citetidyverse. Table showing dataset attained through PDF can be seen in Table \@ref(tab:chuck). In each in gender, whether they used no media, read newspaper weekly, watched television weekly, listened to radio weekly, or used all three was collected and distributions of each variable is shown in Table \@ref(tab:summ). The amount of women and men in each region was recorded, represented as numer_f and numberl_m,  in which it should be noted that women were interviewed at a much larger capacity. This data represents percentages, making comparison of gender classes possible.

@@

It is evident that there are gender influences on access to media as seen through Table \@ref(tab:summ). In the female category, the largest mean amongst variables is those with no access to media. This case is not observed in males as, listening to radio weekly is the largest mean seen. Access to all three media is observed to be the smallest mean in both genders, though men still occupy twice the mean of their female counterpart. Interested on how access to all three media and no mass media compared, Figures \@ref(fig:withoutmed) and \@ref(fig:withmedia)  were made. Figures \@ref(fig:withoutmed) visualizes lack of a mass media source and its distribution among individual regions of Tanznia. It is further shown that females, across all regions have, on average, less access to media compared to males. When comparing to Figure \@ref(fig:withmedia)  representing amount of those with access to all three media, results show that other than Dares Salaam, an urban city, all regions have less access to all media than those with no access to media. These results indicate lack of resources to economically support all three sources.

@@

When availability of resources increase, we were interested to know individuals pattern in mass media collection.  Just looking at male individuals, Figure \@ref(fig:withvsm) demonstrates that when males do not have access to all three media, when x axis is 0, they will listen to radio weekly the most. In fact, the other two sources of mass media, are almost not explored by anyone in this category. There is strong positive linear relationship in having all three media, watching tv and reading the newspaper. There is a weak, almost no relationship seen, however between listening to radio weekly and having ample access to mass media. This indicates that as more resources are available to spend on more mass media, these other forms of communication, other than radio, are then explored. It is possible that they are seen not as a necessity to get information but more of a treat. Figure \@ref(fig:withvsf) displaying the same information for female observes a different trend. All three variables have a positive correlation with using all three methods of media, though more females will listen to the radio more as their access to media increases.

@@

@@

Though there were discrepancies between gender comparison, overall, radio was the most common source of mass media used by those that had used at least one form of media. Radio has been used as a powerful communication tool that has proven to be the most effective media in stimulating agriculture and the development in the rural areas (@bernard2014access). In Tanzania, agriculture plays a huge role in their mixed economy, contributing the largest shared of any sector to the Gross Domestic Product (GDP)(@tanzania). Out of the sources of media studied, radio and television would serve the most useful sources in for farmers. A study found they in fact, out of the sources of media studied, radio and television would serve the most useful sources of agricultural information to farmers in addition to notifying farmers of new agricultural innovations and emergencies(@bernard2014access). Low levels of television access were observed, in both males and females when only one source could be accessed.  Though as resources increased in male participants, watching television was the next desired source of media. In terms of affordability, access to a television would require a high source of income as it requires an adequate source of power. It is understandable why radio is most used as it required less electricity and thus is capable of spreading messages to audiences that just have a receiver with a sufficient power. Print, however, though expensive and majority privately owned, would be more affordable than television. Benefits to agricultural information, which in turn leads to a higher economy, may explain why television was this invested in before reading the newspaper, in males, the main worker population.

@@

Newspaper though slightly more popular in the female demographic, was not a heavily sought-after source of media. To be able to effectively get information out of a newspaper, the audience must be able to read and understand its inputs. Tanzania is known, among east African countries, to have a higher literacy rate a literacy rate of 80%(@spurk2017need), nevertheless, it rural communities tend to have high poverty levels and thus lower literacy rates. The population of Tanzania is predominantly rural, with the proportion of urban residents increasing from 6 percent in 1967 to 18 percent in 1988(@tanzania). Without proper educational resources to understand the newspaper, it remained utilized as a media choice in a small portion of participants. High illiteracy levels, poverty, lack of electricity, mixed with the high cost of information and communication technologies, resulted in Tanzania’s rural areas constrained to a poor and unreliable communication infrastructure(@ogessa2017newspaper). This idea is represented in the data as majority of participants did not use any source of media. Though mass media can provide opportunities in developing countries, the lack of actual use of mass media nulls possible effect as a whole.

@@

Data collection mainly represents data on women. This is despite the large emphasis that mass media has on the agriculture demographic, which is mostly male. With the low sample size selected within this groups, external reliability of results is questionable. Implication of this study may be biased towards the female population, however independent analyse on male data provided is possible in the future. It has been established which forms of media are popular, along with the overall use of media in the Tanzanian population.

The results of this study imply that mass media must be more affordable in order to be optimized. As radio play is already popular and efficient, developing more accessible radio communication to the public could increase effectiveness of mass media.  This because even though radios are the cheapest option, there are still a multitude of individuals that cannot afford to have one in their home. By providing radio play in a public setting such as a radio station, would alleviate an individual purchasing barrier, resulting in benefits of mass media to be seen. It is possible that further development has occurred since data collection in 1996.

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No the dataset will not be distributed to third parties outside of the entity 
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset will be distributed throguh Github
3. *When will the dataset be distributed?*
	- The dataset be distributed in Aprl 2022
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Dataset will not ne  distributed under a copyright or other intellectual property (IP) license 
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- Third parties have not imposed IP-based or other restrictions on the data associated with the instances
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or other regulatory restrictions apply to the dataset or to individual instances
7. *Any other comments?*
	- NA

>>>>

---
title: "The Lack of Accecibility to Media Outlets in Tanzinia in 1996"
subtitle: "May Explain Absence of Benefits Usually Seen Through Mass Media"
author: 
  - Oluwabusayomi Adekuajo
thanks: "Code and data are available at: https://github.com/rubyzero10/Analysis_of_Mass_Media.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The significance of communication on the development of a country cannot be overestimated as the ability to spread information to a large audience supports community and its progression. Lack of progress in developing countries raises the concern of whether beneficial effects of mass media can be observed.  In this paper the prevalence of mass media outlets in Tanzania in 1996 was determined, in which most individuals, in rural ares, did not have access to any avaliable media sources. Our findings have implications to develop affordable access to media through public spaces."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

@@

Communication, more than ever, has great significance to civilisation, so much so that is necessary for survival. The constant development of technology has led to the progression of media to what is known as mass media. Its key function is to communicate various messages to a large audience; common outlets include radio, newspapers, television, internet, social media, and more. The ability to bring information to the masses concerns the successiveness of a society, as some communities have the resources, such as easy access to internet, to take advantage of these communication outlets more than others that struggle to obtain the same source. As the strength of cooperation, obtained through utilization of mass media, is necessary for the development of a community, after the late 90s, new print, radio, and TV enterprises were established, from commercial media outlets to church and community-based stations, in east African developing countries including Tanzania(@spurk2017need). The benefits of this new media was predicted to promote democratic support and enhance accountability and participation of the public sector (@spurk2017need).

Proper implementation of mass media was expected to have led to numerous positive outcomes, however, fulfilment of such expectation had yet to be met, despite its investment in east Africa. In this paper, to determine that people are taking advantage of available mass media, we will analyse the proportion of mass media in the different regions of Tanzania. We compare the prevalence of all media outlets, along with discover possible preferences of media outlets. Media use was further compared by gender to visualize possible influence of gender roles on access to media. With this information we tried to determine circumstances in which benefits of media could be seen in developing countries, specifically Tanzania. Findings show that there was a large percentage of both men and women, in multiples regions of Tanzania, that did not use any forms of mass media. Out of those that did use any or all forms of mass media, the most common use was radio, when compared to reading the newspaper or watching television. Findings are in line with past research that determined that the development of radio in Tanzania has been far more rapid than that of the press, with the potential of reaching the whole country (@probs). Radio is the most common source of media in developing countries due to its affordability and use of less electricity. Thus our finings imply that in order to achieve the right environment to support the development of media, economic viability must be considered and optimized.

The remainder of the paper is split into four sections. Section \@ref(data) explains the data source and its survey collection methodology, potential strengths and weaknesses, as well as our selected data’s characteristics. Section \@ref(results) presents the findings from our methodology. Section \@ref(discussion) provides discussion of results, weaknesses as well as next steps to expand on the findings of the report. Section \@ref(datasheet) includes the data sheet for the dataset put together.

@@

The 1996 Tanzania Demographic and Health Survey is a nationally representative survey with the main purpose to provide information on fertility, family planning, infant and child mortality, maternal and child health and nutrition, knowledge and attitudes of AIDS, and female circumcision(@tanzania). The TDHS is the third national samples survey of it its kind, practicing surveys since 1991. The general objective of the 1996 TDHS, which concerned the data used, is collection of national-level data that will allow the calculation of demographic rates, particularly fertility and childhood mortality rates(@tanzania). The purpose of the TDHS was to provide estimates to the whole country, including rural and urban areas.

@@

After the sampling process described above, 8900 households were selected in which 8141 were occupied as the target sample size. 7969 individuals were interviewed showing a response rate of 98 percent. Out of the eligible 8900 households, 8501 was the target population for females age 15- 49, in which 96 percent were actually interviewed. In contracts, only 2658 was the target population for men age 15-59, from that  85 percent were interviewed. The TDHS thus collected data from 8,120 women aged 15-49 and 2,256 men aged 15-59 in Tanzania in 1996.

## Strengths

The field staff conducting the interviews had to undergo intensive training consisting of interviewing techniques, detailed review of questionnaire items, mock interviews and much more, before being selected for the fieldwork. A series of assessments were given and graded from selected interviews and those who showed extra knowledge with the ability to detect error in completed questionnaires were later chosen to be editors. This training period consisted over the course of three weeks.

@@

In urban areas one member households were more common, as they were more likely to lock up their houses for safety. This made interviewing more difficult and response rate lower in such areas. Response rate was also lower among men despite the larger range in age and lower target sample size. This was observed due to the frequent absence of men seen in households at time of interview, due to work status. This fact must be considered when comparing results between men and women.

@@

Interested in understanding the prevalence of mass media outlets in Tanzania households, data showing percentage of women and men who usually read the newspaper once a week, watch television once a week or listen to radio weekly in Tanzania 1996 was selected. The full table can be seen in Figure \@ref(fig:image), in which specifically region data was selected for the purposes of this paper. Data was gathered as a PDF and was read and manipulated through the use of pdftools @pdf and tidyverse @citetidyverse. Table showing dataset attained through PDF can be seen in Table \@ref(tab:chuck). In each in gender, whether they used no media, read newspaper weekly, watched television weekly, listened to radio weekly, or used all three was collected and distributions of each variable is shown in Table \@ref(tab:summ). The amount of women and men in each region was recorded, represented as numer_f and numberl_m,  in which it should be noted that women were interviewed at a much larger capacity. This data represents percentages, making comparison of gender classes possible.

@@

It is evident that there are gender influences on access to media as seen through Table \@ref(tab:summ). In the female category, the largest mean amongst variables is those with no access to media. This case is not observed in males as, listening to radio weekly is the largest mean seen. Access to all three media is observed to be the smallest mean in both genders, though men still occupy twice the mean of their female counterpart. Interested on how access to all three media and no mass media compared, Figures \@ref(fig:withoutmed) and \@ref(fig:withmedia)  were made. Figures \@ref(fig:withoutmed) visualizes lack of a mass media source and its distribution among individual regions of Tanznia. It is further shown that females, across all regions have, on average, less access to media compared to males. When comparing to Figure \@ref(fig:withmedia)  representing amount of those with access to all three media, results show that other than Dares Salaam, an urban city, all regions have less access to all media than those with no access to media. These results indicate lack of resources to economically support all three sources in rural households.

@@

@@

When availability of resources increases, we were interested to know individuals pattern in mass media collection.  Just looking at male individuals, Figure \@ref(fig:withvsm) demonstrates that when males do not have access to all three media, when x axis is 0, they will listen to radio weekly the most. In fact, the other two sources of mass media, are almost not explored by anyone in this category. There is strong positive linear relationship in having all three media, watching tv and reading the newspaper. There is a weak, almost no relationship seen, however between listening to radio weekly and having ample access to mass media. This indicates that as more resources are available to spend on more mass media, then other forms of communication, other than radio, are explored in the male demographic. It is possible that they are seen not as a necessity to get information. Figure \@ref(fig:withvsf), displaying the same information for females, observes a different trend. All three variables have a positive correlation with using all three methods of media, though more females will listen to the radio more as their access to media increases. Suggesting that access to media is not a priority, but when attainable, radio is most 
preferred.

@@

Though there were discrepancies between gender comparisons, overall, radio was the most common source of mass media used by those that had at least one form of media. Radio has been used as a powerful communication tool that has proven to be the most effective media in stimulating agriculture and the development in the rural areas (@bernard2014access). In Tanzania, agriculture plays a huge role in their mixed economy, contributing the largest shared of any sector to the Gross Domestic Product (GDP)(@tanzania). Out of the sources of media studied, radio and television  serve the most use for farmers. A study found that in fact, out of the sources of media studied, radio and television would serve the most useful sources of agricultural information to farmers in addition to notifying farmers of new agricultural innovations and emergencies(@bernard2014access). Low levels of television access were observed, in both males and females when only one source could be accessed.  Though as resources increased in male participants, watching television was the next desired source of media. In terms of affordability, access to a television would require a high source of income as it requires an adequate source of power. It is understandable then why radio is most used, as it required less electricity and thus is capable of spreading messages to an audience through a receiver with sufficient power. Print, however, though expensive and majorly privately owned, would be more affordable than television. Benefits to agricultural information, which in turn leads to a higher economy, may explain why television was actually invested in before reading the newspaper, in males, while newspaper for females was next invested in.

@@

Newspaper, though slightly more popular in the female demographic, was not a heavily sought-after source of media. To be able to effectively get information out of a newspaper, the audience must be able to read and understand its inputs. Tanzania is known, among east African countries, to have a higher literacy rate, a literacy rate of 80%(@spurk2017need), nevertheless, its rural communities tend to have high poverty levels and thus lower literacy rates. The population of Tanzania is predominantly rural, with the proportion of urban residents increasing from 6 percent in 1967 to 18 percent in 1988(@tanzania). Without proper educational resources needed to understand the newspaper, it remained utilized as a media choice in a small portion of participants. High illiteracy levels, poverty, lack of electricity, mixed with the high cost of information and communication technologies, resulted in Tanzania’s rural areas constrained to a poor and unreliable communication infrastructure(@ogessa2017newspaper). This idea is represented in the data as majority of participants did not use any source of media. Though mass media can provide opportunities in developing countries, the lack of actual use of mass media nulls possible effects as a whole.

@@

Data collection mainly represents data on women. This is despite the large emphasis that mass media has on the agriculture demographic, which is mostly male. With the low sample size selected within this groups, external reliability of results is questionable. Implication of this study may be biased towards the female population, however independent analyse on male data provided is possible in the future.

It has been established which forms of media are popular, along with the overall use of media in the 1996 Tanzanian population. The results of this study imply that mass media must be more affordable in order to be optimized. As radio play is already popular and efficient, developing more accessible radio communication to the public could increase effectiveness of mass media.  This is because even though radios are the cheapest option, there are still a multitude of individuals that cannot afford to have one in their home. By providing radio play in a public setting such as a radio station, the individual purchasing barrier would alleviate in the presents of information, resulting in benefits of mass media to be seen. It is possible that further development has occurred since data collection in 1996.

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No the dataset will not be distributed to third parties outside of the entity 
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset will be distributed throguh Github
3. *When will the dataset be distributed?*
	- The dataset will be distributed in April 2022
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Dataset will not be  distributed under a copyright or other intellectual property (IP) license 
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- Third parties have not imposed IP-based or other restrictions on the data associated with the instances
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or other regulatory restrictions apply to the dataset or to individual instances
7. *Any other comments?*
	- NA
